### 2014-08-11 Studies/2011_San_Francisco/Report/San_Francisco_Fire.tex
San Francisco: Rename tex file

\noindent Table~\ref{tab:fire_info} shows an overview of the timeline of events. A plan view of the structure and the location where the captain collapsed are shown in Fig.~\ref{fig:simp_geom}, and more details of the fire development are provided in the following sections.

\begin{table}[!ht]
\caption[Abridged NIOSH approximate fire event timeline]{Abridged NIOSH approximate fire event timeline~\cite{NIOSH:Bowyer}}
\begin{tabular}{cl}
\toprule
Incident Time  &  Fire Behavior / Fireground Operation                       \\
{(hh:mm:ss)}   &                                                             \\
\midrule
17:15:00       &  Reported working fire with heavy smoke coming from the     \\ 
               &  front and rear of the structure's attic.                   \\
               &                                                             \\
17:16:00       &  Dispatch for a ``Still'' alarm - two engine companies,     \\
               &  two truck companies, and a battalion chief.                \\
               &                                                             \\
17:20:00       &  Attic windows break - heavy smoke out of front attic       \\
               &  window and heavy smoke and fire out of rear.               \\
               &                                                             \\
17:21:00       &  E123 captain and firefighter on second floor               \\
               &  with 1\sfrac{3}{4}-inch hoseline.                          \\
               &                                                             \\
17:21:30       &  First floor exterior porch door opened.                    \\
               &                                                             \\
17:23:00       &  Heavy fire report in rear stairwell and enclosed porch.    \\
               &  No response from E123 captain or firefighter.              \\
               &                                                             \\
17:24:00       &  Water applied to attic from the rear exterior position.    \\
               &                                                             \\
17:25:00       &  SQ5 makes entry through front of structure.                \\
               &                                                             \\
17:27:00       &  After hearing a Mayday call from SQ5, IC calls a Mayday.   \\
               &                                                             \\
17:29:00       &  E123 captain carried out of structure.                     \\
\bottomrule
\end{tabular}
\label{tab:fire_info}
\end{table}

@@

% \begin{figure}[!ht]
% \includegraphics[width=in]{../Figures/1st_Floor_Metric} \\
% \includegraphics[width=in]{../Figures/2nd_Floor_Metric}
% \caption[Plan view of first (top) and second (bottom) floor of structure.]{Plan view of first (top) and second (bottom) floor of structure. Stairs and key rooms are identified using measurements collected by NIST.}
% \label{fig:simp_geom}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=3in]{../Figures/exterior_alpha} \\
% \includegraphics[width=2.67in]{../Figures/exterior_charlie}
% \caption{Photographs showing the front (left) and rear (right) exterior of the structure after the incident.}
% \label{fig:alpha_ex}
% \end{figure}

@@

The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). Two different ventilation scenarios were considered in this study. The first scenario involves a baseline simulation in which the ventilation times and ventilation areas represent our best understanding of the incident. The second scenario, or alternative simulation, does not include the firefighter opening the rear door to make entry to the structure and the subsequent exterior left wall fire penetration. The times of the ventilation changes are provided in Table~\ref{tab:vents}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, any leakage in the structure would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

An unknown parameter in Table~\ref{tab:vents} that needed to be estimated was the time when the second floor interior/porch door failed (160~s into the simulation). The area of failure was also estimated based on the following observations. The bottom hinge from the door appeared to be closed throughout the fire due to the lack of fire damage and the lack of soot deposited on the interior faces of the hinge and remaining wood, as shown in Fig.~\ref{fig:door_frame}. It appears that the door collapsed and failed partially open. In the simulations, only the upper portion of the door was assumed to fail. While this time and area may not be exact, the takeaway from the analysis in the following sections should be impact of the door's failure on the interior conditions of the structure.

@@

\begin{table}[!ht]
\caption[Timeline of ventilation changes in the simulation]{Timeline of ventilation changes in the simulation}
\begin{tabular}{cccll}
\toprule
Baseline    &  Alternative  &  Ventilation  &  Type         &  Side of         \\
Simulation  &  Simulation   &  Area         &               &  Structure       \\
{[s]}       &  {[s]}        &  [m$^2$]      &               &                  \\
\midrule
0           &  0            &  0.11         &  Window       &  Rear            \\
40          &  40           &  0.08         &  Breach (x2)  &  Front and Rear  \\
45          &  45           &  0.12         &  Breach       &  Left            \\
55          &  55           &  0.12         &  Breach       &  Rear            \\
130         &  N/A$^*$      &  1.41         &  Door         &  Rear            \\
135         &  N/A$^*$      &  0.2          &  Breach       &  Left            \\
160         &  160          &  0.80         &  Door         &  Interior        \\
\bottomrule
\end{tabular}
\footnotesize
\\$^*$These ventilation events were not included in the alternative simulation.
\normalsize
\label{tab:vents}
\end{table}

% \begin{figure}[!ht]
% \includegraphics[width=.80in]{../Figures/door_frame_burn}
% \caption[Post-incident image of rear porch door frame.]{Post-incident image of the second floor porch door frame that failed taken from inside the structure looking toward the porch. Note that the hinge is still intact and there is unburned, ``clean'' wood along the bottom portion of the frame.}
% \label{fig:door_frame}
% \end{figure}

@@

As discussed in Section~\ref{sec:geometry}, the structure has dimensions of 17.15~m (56~ft) by 6.15~m (20~ft) and a total height of 8.56~m (28~ft). To ensure adequately resolved fluid flow in and out of the structure, the computational domain was extended beyond the volume of the structure. The total volume of the computational domain was 24~m (79~ft) by 10~m (33~ft) by 10~m (33~ft). A grid resolution of 10~cm (3.9~in) was used, which required 2.4 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. Figure~\ref{fig:geom_grid} shows the front of the structure rendered in Smokeview with the 10~cm mesh. The domain was divided into 16 equally sized meshes (each containing 150,000 grid cells) that could be processed in parallel, which reduced the amount of required calculation time to approximately 1.5~days. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. Table~\ref{tab:mod_param} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

% \begin{figure}[!ht]
% \includegraphics[width=.49in]{../Figures/smv_exterior_grid} \\
% \includegraphics[width=.488in]{../Figures/smv_exterior_grid_rear}
% \caption{Front (left) and rear (right) of the structure with 10~cm computational mesh.}
% \label{fig:geom_grid}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=.60in]{../Figures/smv_exterior_mesh}
% \caption[Front-right side of the structure within the computational domain.]
% {Front-right side of the structure within the computational domain (with multiple meshes). Each of the 16 meshes is 3~m (9.8~ft) by 10~m (33~ft) by 5~m (16.4~ft).}
% \label{fig:mult_mesh}
% \end{figure}

\begin{table}[!ht]
\caption[Relevant fire model input parameters]{Relevant fire model input parameters}
\begin{tabular}{lll}
\toprule
Parameter                                                  & Description                                   & Discussion                                     \\
\midrule
Simulation Time                                            &  5~min                                        & --                                             \\ [.25cm]
Grid Cell Size                                             &  10~cm                                        & Section~\ref{sec:numerical_mesh}               \\ [.25cm]
Ambient Temperature$^*$                                    &  20~$^{\circ}$C (77~$^{\circ}$F)              & --                                             \\ [.1cm]
\multirow{4}{*}{Reaction: Wood~\cite{SFPE:Tewarson}}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     & \multirow{4}{*}{Section~\ref{sec:fire}}        \\
                                                           &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                                \\
                                                           &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                                \\
                                                           &  Heat of Combustion: 16,400~kJ/kg             &                                                \\ [.25cm]
\multirow{2}{*}{Peak HRR}                                  &  Attic Fire: 9~MW                             & \multirow{2}{*}{Section~\ref{sec:fire}}        \\ 
                                                           &  Porch Fire: 3.3~MW                           &                                                \\ [.25cm]                     
\multirow{3}{*}{Material: Wood~\cite{Incropera:1} }        &  $k$: 0.12~\si{W/(m.K)})                      & \multirow{3}{*}{Section~\ref{sec:materials}}   \\
                                                           &  $\rho$: 510~kg/m$^3$                         &                                                \\
                                                           &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                                \\ [.25cm]
\multirow{3}{*}{Material: Gypsum Board~\cite{WAKILI2007}}  &  $k$: 0.28~\si{W/(m.K)}                       &  \multirow{3}{*}{Section~\ref{sec:materials}}  \\ 
                                                           &  $\rho$: 810~kg/m$^3$                         &                                                \\ 
                                                           &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                                \\
\bottomrule
\end{tabular}
\footnotesize
\\ $^{*}$ Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 7~$^{\circ}$C~\cite{NIOSH:Bowyer}.
\normalsize
\label{tab:mod_param}
\end{table}

@@

To examine the results of the simulations, it is important to link the timelines from the fire scene to the simulation times. Table~\ref{tab:firesim} shows the fireground timeline~\cite{NIOSH:Bowyer} along with the corresponding simulation times.

@@

\\ For specific information regarding ventilation areas used in the FDS model, refer to Table~\ref{tab:vents}.

@@

\label{tab:firesim}

@@

% \begin{figure}[!ht]
% \includegraphics[width=.80in]{../Figures/Chicago_Fire_HRR}
% \caption[HRR versus time from the FDS simulations.]
% {HRR versus time from the FDS simulations. The first vertical line indicates the time at which the back door failed (130~s), and the second vertical line indicates the time at which the interior hallway door failed (160~s).}
% \label{fig:hrr}
% \end{figure}

In Fig.~\ref{fig:hrr}, the solid line represents the baseline FDS simulation (this follows the noted door openings from the incident). The dashed line represents an alternative series of events in which the rear door does not get opened and the subsequent breach of the rear porch wall on the left side of the structure does not occur. The dash-dotted line represents the prescribed (or nominal) HRR that was input into FDS. As shown in Fig.~\ref{fig:hrr}, the HRR output from FDS is less than the prescribed HRR because the fire in the structure becomes ventilation limited. The maximum HRR value is slightly less than 9~MW before the fire becomes ventilation limited and begins to decay. The HRR is reduced to 5~MW as the oxygen inside the structure is consumed. There is an insufficient amount of oxygen in the structure to burn all of the fuel, therefore combustion occurs locally at the vents where fresh air is available. Figure~\ref{fig:smv_ext_fire} shows flaming combustion through the failed window on the rear side of the structure and attic breach just after the second fire ignites at 40~s. After additional wall breaches occur at 45~s and 55~s (see Table~\ref{tab:vents}), the abrupt change in ventilation causes the HRR to increase to 9~MW before it decreases to approximately 6~MW. The HRR then increases towards a sustainable steady-state value.

Figure~\ref{fig:hrr} indicates that the HRRs for both of the ventilation scenarios in FDS are identical until 130~s (HRR $\approx$ 7~MW). At 130~s, the rear door is opened in the baseline simulation. At 135~s, there is an additional breach on the left side of the structure (Table~\ref{tab:vents}). The open door and breach act as additional pressure releases for the structure. Again, the abrupt change in ventilation causes the HRR to increase from approximately 7~MW to 9.7~MW at 139~s. At the time of the door failure (160~s), the HRR in both cases was approximately 9~MW because of the increased amount of ventilation. While there might be some locally different burning characteristics between the baseline simulation and the alternative simulation, the interior conditions in both cases presented a significant hazard to firefighters inside of the structure. Therefore, the analysis will focus on the results of the baseline simulation.

% \begin{figure}[!ht]
% \includegraphics[width=.675in]{../Figures/smv_exterior_fire}
% \caption{Smokeview rendering of exterior combustion occurring on the rear side of the structure.}
% \label{fig:smv_ext_fire}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=.75in]{../Figures/west_50th_baseline_pres2}
% \input{../Scripts/colorbar_metric_pres}
% \caption[FDS simulated pressure contours, 1~s prior to interior door failure.]{FDS simulated pressure contours, 1~s prior to interior door failure.}
% \label{fig:pres_159s}
% \end{figure}

@@

% \begin{landscape}
% \begin{figure}[!ht]
% \includegraphics[width=1.1in]{../Figures/west_50th_baseline_velo_175}
% \input{../Scripts/colorbar_std_velo}
% \input{../Scripts/colorbar_metric_velo}
% \caption{FDS simulated velocity vectors, 15~s after interior door failure.}
% \label{fig:velo_175s}
% \end{figure}
% \end{landscape}
% \newpage

@@

% \begin{figure}[!ht]
% \includegraphics[width=.675in]{../Figures/west_50th_baseline_129}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption{FDS simulated temperature contours, 1~s prior to rear door opening.}
% \label{fig:temp_129s}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=.675in]{../Figures/west_50th_baseline_159}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption{FDS simulated temperature contours, 1~s prior to interior door failure.}
% \label{fig:temp_159s}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=.675in]{../Figures/west_50th_baseline_161}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption{FDS simulated temperature contours, 1~s after interior door failure.}
% \label{fig:temp_161s}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=.675in]{../Figures/west_50th_baseline_175}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption{FDS simulated temperature contours, 15~s after the interior door failure.}
% \label{fig:temp_175s}
% \end{figure}

@@

\section{Simulated Hallway Flow Path}
From on-scene reports (Section~\ref{sec:incident_summary}) and the FDS simulation results (Section~\ref{sec:temperature}), the interior conditions on the second floor of the structure were initially tenable. From the NIOSH report~\cite{NIOSH:Bowyer}, while the IC was inside the structure on the second floor, only light haze and a glow around the door between the hallway and enclosed porch was observed. This indicated the presence of fire in the area of the enclosed porch. The simulation results indicate a high hazard area (temperatures greater than 260~$^{\circ}$C or 500~$^{\circ}$F) in the attic and the enclosed porch at this time. By the time the E123 captain and firefighter arrived on the second floor (Table~\ref{tab:firesim}) and approached the rear of the structure, the rear porch was fully involved. Temperature conditions greater than 260~$^{\circ}$C (500~$^{\circ}$F) were present on the porch side of the steel-faced, wood-framed door at the rear of the structure. Exposure to these conditions likely weakened the structural integrity of the door, which may have led to a partial failure (Table~\ref{tab:firesim}). Figure~\ref{fig:chicago_doorfold} shows the state of the rear door after the fire. Note the large crease in the steel face of the door just above the knob.

% \begin{figure}[!ht]
% \includegraphics[width=.4in]{../Figures/Porch_Door_1}
% \caption{Post-fire image of interior door.}
% \label{fig:chicago_doorfold}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=3in]{../Figures/SC_Door_1} \\
% \includegraphics[width=2in]{../Figures/SC_Door_2}
% \caption{Post-experiment images of door failures.}
% \label{fig:SC_doorfold}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=.7in]{../Figures/ChicagoFlow}
% \caption{Summary of simulated flow path on the second floor after interior door failure.}
% \label{fig:flowpath_1}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=.68in]{../Figures/west_50th_baseline_top_159_6ft}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption[Temperature contours on the second floor, 1~s prior to interior door failure.]
% {Temperature contours on the second floor, 1~s prior to interior door failure, 1.83~m (6~ft) above the floor.}
% \label{fig:temp_top_159}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=.68in]{../Figures/west_50th_baseline_top_165_6ft}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption[Temperature contours on the second floor, 5~s after interior door failure.]
% {Temperature contours on the second floor, 5~s after interior door failure, 1.83~m (6~ft) above the floor.}
% \label{fig:temp_top_165}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=.68in]{../Figures/west_50th_baseline_top_220_6ft}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption[Temperature contours on the second floor, 60~s after interior door failure.]
% {Temperature contours on the second floor, 60~s after interior door failure, 1.83~m (6~ft) above the floor.}
% \label{fig:temp_top_220}
% \end{figure}

>>>>

\noindent Table~\ref{tab:incident_timeline} shows an overview of the timeline of events. A plan view of the structure and the location where the captain collapsed are shown in Fig.~\ref{fig:simp_geom}, and more details of the fire development are provided in the following sections.

\begin{table}[!ht]
\caption[Abridged NIOSH approximate fire event timeline]{Abridged NIOSH approximate fire event timeline~\cite{NIOSH:Bowyer}}
\begin{tabular}{cl}
\toprule
Incident Time  &  Fire Behavior / Fireground Operation                       \\
{(hh:mm:ss)}   &                                                             \\
\midrule
17:15:00       &  Reported working fire with heavy smoke coming from the     \\ 
               &  front and rear of the structure's attic.                   \\
               &                                                             \\
17:16:00       &  Dispatch for a ``Still'' alarm - two engine companies,     \\
               &  two truck companies, and a battalion chief.                \\
               &                                                             \\
17:20:00       &  Attic windows break - heavy smoke out of front attic       \\
               &  window and heavy smoke and fire out of rear.               \\
               &                                                             \\
17:21:00       &  E123 captain and firefighter on second floor               \\
               &  with 1\sfrac{3}{4}-inch hoseline.                          \\
               &                                                             \\
17:21:30       &  First floor exterior porch door opened.                    \\
               &                                                             \\
17:23:00       &  Heavy fire report in rear stairwell and enclosed porch.    \\
               &  No response from E123 captain or firefighter.              \\
               &                                                             \\
17:24:00       &  Water applied to attic from the rear exterior position.    \\
               &                                                             \\
17:25:00       &  SQ5 makes entry through front of structure.                \\
               &                                                             \\
17:27:00       &  After hearing a Mayday call from SQ5, IC calls a Mayday.   \\
               &                                                             \\
17:29:00       &  E123 captain carried out of structure.                     \\
\bottomrule
\end{tabular}
\label{tab:incident_timeline}
\end{table}

@@

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/1st_Floor_Metric} \\
% \includegraphics[width=5in]{../Figures/2nd_Floor_Metric}
% \caption[Plan view of first (top) and second (bottom) floor of structure.]{Plan view of first (top) and second (bottom) floor of structure. Stairs and key rooms are identified using measurements collected by NIST.}
% \label{fig:simp_geom}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/exterior_alpha} \\
% \includegraphics[width=5in]{../Figures/exterior_charlie}
% \caption{Photographs showing the front (left) and rear (right) exterior of the structure after the incident.}
% \label{fig:alpha_ex}
% \end{figure}

@@

The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). Two different ventilation scenarios were considered in this study. The first scenario involves a baseline simulation in which the ventilation times and ventilation areas represent our best understanding of the incident. The second scenario, or alternative simulation, does not include the firefighter opening the rear door to make entry to the structure and the subsequent exterior left wall fire penetration. The times of the ventilation changes are provided in Table~\ref{tab:ventilation_timeline}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, any leakage in the structure would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

An unknown parameter in Table~\ref{tab:ventilation_timeline} that needed to be estimated was the time when the second floor interior/porch door failed (160~s into the simulation). The area of failure was also estimated based on the following observations. The bottom hinge from the door appeared to be closed throughout the fire due to the lack of fire damage and the lack of soot deposited on the interior faces of the hinge and remaining wood, as shown in Fig.~\ref{fig:door_frame}. It appears that the door collapsed and failed partially open. In the simulations, only the upper portion of the door was assumed to fail. While this time and area may not be exact, the takeaway from the analysis in the following sections should be impact of the door's failure on the interior conditions of the structure.

@@

\begin{table}[!ht]
\caption[Timeline of ventilation changes in the simulation]{Timeline of ventilation changes in the simulation}
\begin{tabular}{cccll}
\toprule
Baseline    &  Alternative  &  Ventilation  &  Type         &  Side of         \\
Simulation  &  Simulation   &  Area         &               &  Structure       \\
{[s]}       &  {[s]}        &  [m$^2$]      &               &                  \\
\midrule
0           &  0            &  0.11         &  Window       &  Rear            \\
40          &  40           &  0.08         &  Breach (x2)  &  Front and Rear  \\
45          &  45           &  0.12         &  Breach       &  Left            \\
55          &  55           &  0.12         &  Breach       &  Rear            \\
130         &  N/A$^*$      &  1.41         &  Door         &  Rear            \\
135         &  N/A$^*$      &  0.2          &  Breach       &  Left            \\
160         &  160          &  0.80         &  Door         &  Interior        \\
\bottomrule
\end{tabular}
\footnotesize
\\$^*$These ventilation events were not included in the alternative simulation.
\normalsize
\label{tab:ventilation_timeline}
\end{table}

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/door_frame_burn}
% \caption[Post-incident image of rear porch door frame.]{Post-incident image of the second floor porch door frame that failed taken from inside the structure looking toward the porch. Note that the hinge is still intact and there is unburned, ``clean'' wood along the bottom portion of the frame.}
% \label{fig:door_frame}
% \end{figure}

@@

As discussed in Section~\ref{sec:geometry}, the structure has dimensions of 17.15~m (56~ft) by 6.15~m (20~ft) and a total height of 8.56~m (28~ft). To ensure adequately resolved fluid flow in and out of the structure, the computational domain was extended beyond the volume of the structure. The total volume of the computational domain was 24~m (79~ft) by 10~m (33~ft) by 10~m (33~ft). A grid resolution of 10~cm (3.9~in) was used, which required 2.4 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. Figure~\ref{fig:geom_grid} shows the front of the structure rendered in Smokeview with the 10~cm mesh. The domain was divided into 16 equally sized meshes (each containing 150,000 grid cells) that could be processed in parallel, which reduced the amount of required calculation time to approximately 1.5~days. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. Table~\ref{tab:model_parameters} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/smv_exterior_grid} \\
% \includegraphics[width=5in]{../Figures/smv_exterior_grid_rear}
% \caption{Front (left) and rear (right) of the structure with 10~cm computational mesh.}
% \label{fig:geom_grid}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/smv_exterior_mesh}
% \caption[Front-right side of the structure within the computational domain.]
% {Front-right side of the structure within the computational domain (with multiple meshes). Each of the 16 meshes is 3~m (9.8~ft) by 10~m (33~ft) by 5~m (16.4~ft).}
% \label{fig:mult_mesh}
% \end{figure}

\begin{table}[!ht]
\caption[Relevant fire model input parameters]{Relevant fire model input parameters}
\begin{tabular}{lll}
\toprule
Parameter                                 &  Description                                  &  Discussion                        \\
\midrule
Simulation Time                           &  5~min                                        &  --                                \\
Grid Cell Size                            &  10~cm                                        &  Section~\ref{sec:numerical_mesh}  \\
Ambient Temperature$^*$                   &  20~$^{\circ}$C (77~$^{\circ}$F)              &  --                                \\
Reaction: Wood~\cite{SFPE:Tewarson}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     &  Section~\ref{sec:fire}            \\
                                          &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                    \\
                                          &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                    \\
                                          &  Heat of Combustion: 16,400~kJ/kg             &                                    \\
Peak HRR                                  &  Attic Fire: 9~MW                             &  Section~\ref{sec:fire}            \\
                                          &  Porch Fire: 3.3~MW                           &                                    \\
Material: Wood~\cite{Incropera:1}         &  $k$: 0.12~\si{W/(m.K)})                      &  Section~\ref{sec:materials}       \\
                                          &  $\rho$: 510~kg/m$^3$                         &                                    \\
                                          &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                    \\
Material: Gypsum Board~\cite{WAKILI2007}  &  $k$: 0.28~\si{W/(m.K)}                       &  Section~\ref{sec:materials}       \\
                                          &  $\rho$: 810~kg/m$^3$                         &                                    \\
                                          &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                    \\
\bottomrule
\end{tabular}
\footnotesize
\\ $^{*}$ Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 7~$^{\circ}$C~\cite{NIOSH:Bowyer}.
\normalsize
\label{tab:model_parameters}
\end{table}

@@

To examine the results of the simulations, it is important to link the timelines from the fire scene to the simulation times. Table~\ref{tab:combined_timeline} shows the fireground timeline~\cite{NIOSH:Bowyer} along with the corresponding simulation times.

@@

\\ For specific information regarding ventilation areas used in the FDS model, refer to Table~\ref{tab:ventilation_timeline}.

@@

\label{tab:combined_timeline}

@@

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/Chicago_Fire_HRR}
% \caption[HRR versus time from the FDS simulations.]
% {HRR versus time from the FDS simulations. The first vertical line indicates the time at which the back door failed (130~s), and the second vertical line indicates the time at which the interior hallway door failed (160~s).}
% \label{fig:hrr}
% \end{figure}

In Fig.~\ref{fig:hrr}, the solid line represents the baseline FDS simulation (this follows the noted door openings from the incident). The dashed line represents an alternative series of events in which the rear door does not get opened and the subsequent breach of the rear porch wall on the left side of the structure does not occur. The dash-dotted line represents the prescribed (or nominal) HRR that was input into FDS. As shown in Fig.~\ref{fig:hrr}, the HRR output from FDS is less than the prescribed HRR because the fire in the structure becomes ventilation limited. The maximum HRR value is slightly less than 9~MW before the fire becomes ventilation limited and begins to decay. The HRR is reduced to 5~MW as the oxygen inside the structure is consumed. There is an insufficient amount of oxygen in the structure to burn all of the fuel, therefore combustion occurs locally at the vents where fresh air is available. Figure~\ref{fig:smv_ext_fire} shows flaming combustion through the failed window on the rear side of the structure and attic breach just after the second fire ignites at 40~s. After additional wall breaches occur at 45~s and 55~s (see Table~\ref{tab:ventilation_timeline}), the abrupt change in ventilation causes the HRR to increase to 9~MW before it decreases to approximately 6~MW. The HRR then increases towards a sustainable steady-state value.

Figure~\ref{fig:hrr} indicates that the HRRs for both of the ventilation scenarios in FDS are identical until 130~s (HRR $\approx$ 7~MW). At 130~s, the rear door is opened in the baseline simulation. At 135~s, there is an additional breach on the left side of the structure (Table~\ref{tab:ventilation_timeline}). The open door and breach act as additional pressure releases for the structure. Again, the abrupt change in ventilation causes the HRR to increase from approximately 7~MW to 9.7~MW at 139~s. At the time of the door failure (160~s), the HRR in both cases was approximately 9~MW because of the increased amount of ventilation. While there might be some locally different burning characteristics between the baseline simulation and the alternative simulation, the interior conditions in both cases presented a significant hazard to firefighters inside of the structure. Therefore, the analysis will focus on the results of the baseline simulation.

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/smv_exterior_fire}
% \caption{Smokeview rendering of exterior combustion occurring on the rear side of the structure.}
% \label{fig:smv_ext_fire}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_pres2}
% \input{../Scripts/colorbar_metric_pres}
% \caption[FDS simulated pressure contours, 1~s prior to interior door failure.]{FDS simulated pressure contours, 1~s prior to interior door failure.}
% \label{fig:pres_159s}
% \end{figure}

@@

% \begin{landscape}
% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_velo_175}
% \input{../Scripts/colorbar_std_velo}
% \input{../Scripts/colorbar_metric_velo}
% \caption{FDS simulated velocity vectors, 15~s after interior door failure.}
% \label{fig:velo_175s}
% \end{figure}
% \end{landscape}
% \newpage

@@

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_129}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption{FDS simulated temperature contours, 1~s prior to rear door opening.}
% \label{fig:temp_129s}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_159}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption{FDS simulated temperature contours, 1~s prior to interior door failure.}
% \label{fig:temp_159s}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_161}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption{FDS simulated temperature contours, 1~s after interior door failure.}
% \label{fig:temp_161s}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_175}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption{FDS simulated temperature contours, 15~s after the interior door failure.}
% \label{fig:temp_175s}
% \end{figure}

@@

\section{Simulated Hallway Flow Path}
From on-scene reports (Section~\ref{sec:incident_summary}) and the FDS simulation results (Section~\ref{sec:temperature}), the interior conditions on the second floor of the structure were initially tenable. From the NIOSH report~\cite{NIOSH:Bowyer}, while the IC was inside the structure on the second floor, only light haze and a glow around the door between the hallway and enclosed porch was observed. This indicated the presence of fire in the area of the enclosed porch. The simulation results indicate a high hazard area (temperatures greater than 260~$^{\circ}$C or 500~$^{\circ}$F) in the attic and the enclosed porch at this time. By the time the E123 captain and firefighter arrived on the second floor (Table~\ref{tab:combined_timeline}) and approached the rear of the structure, the rear porch was fully involved. Temperature conditions greater than 260~$^{\circ}$C (500~$^{\circ}$F) were present on the porch side of the steel-faced, wood-framed door at the rear of the structure. Exposure to these conditions likely weakened the structural integrity of the door, which may have led to a partial failure (Table~\ref{tab:combined_timeline}). Figure~\ref{fig:chicago_doorfold} shows the state of the rear door after the fire. Note the large crease in the steel face of the door just above the knob.

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/Porch_Door_1}
% \caption{Post-fire image of interior door.}
% \label{fig:chicago_doorfold}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/SC_Door_1} \\
% \includegraphics[width=5in]{../Figures/SC_Door_2}
% \caption{Post-experiment images of door failures.}
% \label{fig:SC_doorfold}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/ChicagoFlow}
% \caption{Summary of simulated flow path on the second floor after interior door failure.}
% \label{fig:flowpath_1}
% \end{figure}

@@

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_top_159_6ft}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption[Temperature contours on the second floor, 1~s prior to interior door failure.]
% {Temperature contours on the second floor, 1~s prior to interior door failure, 1.83~m (6~ft) above the floor.}
% \label{fig:temp_top_159}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_top_165_6ft}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption[Temperature contours on the second floor, 5~s after interior door failure.]
% {Temperature contours on the second floor, 5~s after interior door failure, 1.83~m (6~ft) above the floor.}
% \label{fig:temp_top_165}
% \end{figure}

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/west_50th_baseline_top_220_6ft}
% \input{../Scripts/colorbar_std} 
% \input{../Scripts/colorbar_metric}
% \caption[Temperature contours on the second floor, 60~s after interior door failure.]
% {Temperature contours on the second floor, 60~s after interior door failure, 1.83~m (6~ft) above the floor.}
% \label{fig:temp_top_220}
% \end{figure}
